---
title: "DATA 606 Project Proposal"
author: "Nathan Cooper"
date: "October 20, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding:  hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, eval=TRUE}
suppressMessages(suppressWarnings(library(tidyr)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(psych)))
suppressMessages(suppressWarnings(library(stringr)))
```

### Data Preparation

The data are sourced from 538's github page:

```{r data_prep, eval=TRUE}

# I pull the data directly from the website and organize it by Major
#category for easier subsetting
url1 <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv"
all_ages <- url1 %>% read.csv(stringsAsFactors = FALSE) %>% tbl_df() %>% arrange(Major_category)

#Below I subset all the data by Major Category I only print 1 table
#to make a cleaner presentation.
all_ages_ag <- all_ages %>% filter(Major_category == "Agriculture & Natural Resources")
head(all_ages_ag)

all_ages_art <- all_ages %>% filter(Major_category == "Arts")

all_ages_bio <- all_ages %>% filter(Major_category == "Biology & Life Science")

all_ages_bsn <- all_ages %>% filter(Major_category == "Business")

all_ages_cj <- all_ages %>% filter(Major_category == "Communications & Journalism")


all_ages_com <- all_ages %>% filter(Major_category == "Computers & Mathematics")

all_ages_ed <- all_ages %>% filter(Major_category == "Education")


all_ages_eng <- all_ages %>% filter(Major_category == "Engineering")

all_ages_hlt <- all_ages %>% filter(Major_category == "Health")

all_ages_la <- all_ages %>% filter(Major_category == "Humanities & Liberal Arts" | Major_category == "Interdisciplinary")

all_ages_ia <- all_ages %>% filter(Major_category == "Industrial Arts & Consumer Services")

all_ages_law <- all_ages %>% filter(Major_category == "Law & Public Policy")

all_ages_sci <- all_ages %>% filter(Major_category == "Physical Sciences")

all_ages_psy <- all_ages %>% filter(Major_category == "Psychology & Social Work")

all_ages_ssc <- all_ages %>% filter(Major_category == "Social Science")

# I repeat the process for graduate students, again only printing
#1 table
url2 <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/grad-students.csv"
grad_stdnt <- url2 %>% read.csv(stringsAsFactors = FALSE) %>% tbl_df() %>% arrange(Major_category)

#Subsets
grad_ag <- grad_stdnt %>% filter(Major_category == "Agriculture & Natural Resources")
head(grad_ag)

grad_art <- grad_stdnt %>% filter(Major_category == "Arts")

grad_bio <- grad_stdnt %>% filter(Major_category == "Biology & Life Science")

grad_bsn <- grad_stdnt %>% filter(Major_category == "Business")

grad_cj <- grad_stdnt %>% filter(Major_category == "Communications & Journalism")

grad_com <- grad_stdnt %>% filter(Major_category == "Computers & Mathematics")

grad_ed <- grad_stdnt %>% filter(Major_category == "Education")

grad_eng <- grad_stdnt %>% filter(Major_category == "Engineering")

grad_hlt <- grad_stdnt %>% filter(Major_category == "Health")

grad_la <- grad_stdnt %>% filter(Major_category == "Humanities & Liberal Arts" | Major_category == "Interdisciplinary")

grad_ia <- grad_stdnt %>% filter(Major_category == "Industrial Arts & Consumer Services")

grad_law <- grad_stdnt %>% filter(Major_category == "Law & Public Policy")

grad_sci <- grad_stdnt %>% filter(Major_category == "Physical Sciences")

grad_psy <- grad_stdnt %>% filter(Major_category == "Psychology & Social Work")

grad_ssc <- grad_stdnt %>% filter(Major_category == "Social Science")



#Repeat for only undergraduates 
url3 <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv"
rct_grad <- url3 %>% read.csv(stringsAsFactors = FALSE) %>% tbl_df() %>% arrange(Major_category)

rct_ag <- rct_grad %>% filter(Major_category == "Agriculture & Natural Resources")
head(rct_ag)

rct_art <- rct_grad %>% filter(Major_category == "Arts")

rct_bio <- rct_grad %>% filter(Major_category == "Biology & Life Science")

rct_bsn <- rct_grad %>% filter(Major_category == "Business")

rct_cj <- rct_grad %>% filter(Major_category == "Communications & Journalism")

rct_com <- rct_grad %>% filter(Major_category == "Computers & Mathematics")

rct_ed <- rct_grad %>% filter(Major_category == "Education")

rct_eng <- rct_grad %>% filter(Major_category == "Engineering")

rct_hlt <- rct_grad %>% filter(Major_category == "Health")

rct_la <- rct_grad %>% filter(Major_category == "Humanities & Liberal Arts" | Major_category == "Interdisciplinary")

rct_ia <- rct_grad %>% filter(Major_category == "Industrial Arts & Consumer Services")

rct_law <- rct_grad %>% filter(Major_category == "Law & Public Policy")

rct_sci <- rct_grad %>% filter(Major_category == "Physical Sciences")

rct_psy <- rct_grad %>% filter(Major_category == "Psychology & Social Work")

rct_ssc <- rct_grad %>% filter(Major_category == "Social Science")
```

### Research Question

Which college majors offer the best opportunities in terms of unemployment rate and salary?

### Cases

#### All_ages

Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include both undergrads and grad students.

#### Grad Students

Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include only grad students aged 25+ years.

#### Recent Grads

Each case represents majors offered by colleges and universities in the US. There are 173 majors represented. These data include only undergraduate students aged <28 years. These data also include gender statistics.

#### Data Collection

These Data were collated by the 538 website: http://www.fivethirtyeight.com and was posted to their github page: https://github.com/fivethirtyeight/data/tree/master/college-majors. They in turn used data from:

 "All data is from American Community Survey 2010-2012 Public Use Microdata Series.

Download data here: http://www.census.gov/programs-surveys/acs/data/pums.html

Documentation here: http://www.census.gov/programs-surveys/acs/technical-documentation/pums.html

Major categories are from Carnevale et al, "What's It Worth?: The Economic Value of College Majors." Georgetown University Center on Education and the Workforce, 2011. http://cew.georgetown.edu/whatsitworth"

#### Type of Study

This is an observational Study

#### Response Variable

The response variable are the college majors and are categorical.

#### Explanatory Variables

The explanatory variables are the counts of employed and unemployed college degree holders and the statistics of their income. These data are numerical.

### Relavent Summary Statistics

First we will look at overall unemployment rate for the 3 categories: all ages, recent grads, and grad students.
```{r sum_stats, eval=TRUE}
summary(all_ages$Unemployment_rate)
summary(rct_grad$Unemployment_rate)
summary(grad_stdnt$Grad_unemployment_rate)
unempl <- cbind(all_ages$Unemployment_rate, rct_grad$Unemployment_rate, grad_stdnt$Grad_unemployment_rate)
barplot(unempl/nrow(unempl), names.arg = c("All", "Recent Grad", "Grad Student"), xlab = "Unemployment Rate", col = rainbow(nrow(unempl)))
```

It appears that people holding only a Bachelor's degree have nearly twice as high median unemployment as those with higher degrees.

We will also look at median income for the three categories.

```{r med_inc, eval=TRUE}
summary(all_ages$Median)
hist(all_ages$Median, main = "Histogram for Median Income All Ages", xlab = "Median Income by Major All Ages (USD)", col = "dark blue")
summary(rct_grad$Median)
hist(rct_grad$Median, main = "Histogram for Median Income Recent Grads", xlab = "Median Income by Major Recent Grads (USD)", col = "dark blue")
summary(grad_stdnt$Grad_median)
hist(grad_stdnt$Grad_median, main = "Histogram for Median Income Grad Students", xlab = "Median Income by Major Grad Student (USD)", col = "dark blue")
medsal <- cbind(all_ages$Median, rct_grad$Median, grad_stdnt$Grad_median)
barplot(medsal/nrow(medsal), names.arg = c("All", "Recent Grad", "Grad Student"), xlab = "Median Salary", col = rainbow(nrow(medsal)))
```

