---
title: "DATA 607 MySQL to MongoDB"
author: "Nathan Cooper"
date: "November 15, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MySQL to CSV

First we will take the Flights Database from MySQL and transfer it to a .csv file with the help of https://www.r-bloggers.com/dump-mysql-to-csv-using-r/

```{r sql-csv, eval=TRUE}
suppressMessages(suppressWarnings(library(RMySQL)))
db <- dbConnect(MySQL(), user='*******', password='*********', dbname='flights', host='localhost')
tables <- dbListTables(db)
tables
for (i in 1 : length(tables)){
  temp<-(dbReadTable(db, tables[i]))
  write.table(temp, tables[i], row.names=F, sep = ',')
}
flights <- data.frame(read.csv("~\\Dataset\\flights", stringsAsFactors = FALSE))
head(flights)
ontime <- data.frame(read.csv("~\\Dataset\\ontime", stringsAsFactors = FALSE))
head(ontime)
planes <- data.frame(read.csv("~\\Dataset\\planes", stringsAsFactors = FALSE))
head(planes)
weather <- data.frame(read.csv("~\\Dataset\\weather", stringsAsFactors = FALSE))
head(weather)
airlines <- data.frame(read.csv("~\\Dataset\\airlines", stringsAsFactors = FALSE))
head(airlines)
airports <- data.frame(read.csv("~\\Dataset\\airports", stringsAsFactors = FALSE))
head(airports)
cities <- data.frame(read.csv("~\\Dataset\\cities", stringsAsFactors = FALSE))
head(cities)
```

## From dataframe to MongoDB

Now we will use the dataframes to populate the MongoDB.

```{r mngdb, eval=TRUE}
suppressMessages(suppressWarnings(library(rmongodb)))
mongo = mongo.create(host = "localhost")
mongo.is.connected(mongo)
b1=mongo.bson.from.df(flights)
b2=mongo.bson.from.df(ontime)
b3=mongo.bson.from.df(planes)
b4=mongo.bson.from.df(weather)
b5=mongo.bson.from.df(airlines)
b6=mongo.bson.from.df(airports)
b7=mongo.bson.from.df(cities)
b1[[1]]
b2[[1]]
b3[[1]]
b4[[1]]
b5[[1]]
b6[[1]]
b7[[1]]
mongo.insert(mongo,"flights",b1)
mongo.insert(mongo,"ontime",b2)
mongo.insert(mongo,"planes",b3)
mongo.insert(mongo,"weather",b4)
mongo.insert(mongo,"airlines",b5)
mongo.insert(mongo,"airports",b6)
mongo.insert(mongo,"cities",b7)
```

## From csv file to Neo4J

First I established a database called 'Flights' using the Neo4J desktop app. Neo4J assigned this database to an http port of 11001. To access this database in Chrome, I entered http:\\localhost:11001 note this is different than the default port of 7474.

To use the 'LOAD CSV WITH HEADERS' cypher command I had to move the csv files into Neo4J desktop import file or the following warning would appear "Couldn't load the external resource at: file:/C:/Users/Nate/AppData/Roaming/Neo4j%20Desktop/Application/neo4jDatabases/database-c2becb82-5c85-430d-bb42-4ad6749f1f0b/current/import/airlines"

The successfull Cypher queries were as follows:

> load csv with headers from "file:///airlines" as airlines create (n:Airlines {carrier: airlines.carrier, name: airlines.name})

> MATCH (n:Airlines) RETURN n LIMIT 1

returned: 
\{
  "name": "Endeavor Air Inc.\r",
  "carrier": "9E"
\}

> load csv with headers from "file:///airports" as airports create (n:Airports {faa: airports.faa, name: airports.name, lat:	airports.lat, lon:	airports.lon, alt:	airports.alt, tz:	airports.tz, dst:	airports.dst})

> MATCH (n:Airports) RETURN n LIMIT 1

returned: 

{
  "dst": "A",
  "tz": "-5",
  "faa": "04G",
  "alt": "1044",
  "name": "Lansdowne Airport",
  "lon": "-80.6195833",
  "lat": "41.1304722"
}

> load csv with headers from "file:///planes" as planes create (n:Planes {tailnum: planes.tailnum, year:	planes.year, type:	planes.type, manufacturer:	planes.manufacturer, model:	planes.model,engines:	planes.engines, seats: planes.seats,	speed: planes.speed.	engine: planes.engine})

> MATCH (n:Planes) RETURN n LIMIT 1

returned:

{
  "tailnum": "N10156",
  "year": "2004",
  "engine": "Turbo-fan",
  "engines": "2",
  "model": "EMB-145XR",
  "type": "Fixed wing multi engine",
  "seats": "55",
  "speed": "NA",
  "manufacturer": "EMBRAER"
}

> load csv with headers from "file:///weather" as weather create (n:Weather {origin:weather.origin, year:weather.year, month:weather.month,	day:weather.day,	hour:weather.hour,	temp:weather.temp,	dewp:weather.dewp,	humid:weather.humid,	wind_dir:weather.wind_dir,	wind_speed:weather.wind_speed,	wind_gust:weather.wind_gust,	precip:weather.precip,	pressure:weather.pressure,	visib:weather.visib})

> MATCH (n:Weather) RETURN n LIMIT 1

returned:

{
  "temp": "37.04",
  "year": "2013",
  "origin": "EWR",
  "visib": "10",
  "humid": "53.97",
  "wind_dir": "230",
  "pressure": "1013.9",
  "wind_gust": "11.9186514756",
  "precip": "0",
  "month": "1",
  "dewp": "21.92",
  "hour": "0",
  "wind_speed": "10.35702",
  "day": "1"
}

> load csv with headers from "file:///flights" as flights create (n:Flights {year:flights.year,	month:flights.month,	day:flights.day,	dep_time:flights.dep_time,	dep_delay:flights.dep_delay,	arr_time:flights.arr_time,	arr_delay:flights.arr_delay,	carrier:flights.carrier,	tailnum:flights.tailnum,	flight:flights.flight,	origin:flights.origin,	dest:flights.dest,	air_time:flights.air_time,	distance:flights.distance,	hour:flights.hour,	minute:flights.minute})

Note the retured values once completed:
Added 336776 labels, created 336776 nodes, set 5388416 properties, completed after 14772 ms.

This required me to allot 8 Gb of RAM to Neo4J to complete.

> MATCH (n:Flights) RETURN n LIMIT 1

{
  "flight": "1545",
  "arr_delay": "11",
  "distance": "1400",
  "tailnum": "N14228",
  "year": "2013",
  "origin": "EWR",
  "dep_time": "517",
  "dep_delay": "2",
  "dest": "IAH",
  "minute": "17",
  "carrier": "UA",
  "month": "1",
  "hour": "5",
  "arr_time": "830",
  "air_time": "227",
  "day": "1"
}

These were from the Alsaka Air vs. AM West assignment:

> load csv with headers from "file:///cities" as cities create (n:Cities {city_id:cities.city_id,	city:cities.city})

> MATCH (n:Cities) RETURN n LIMIT 1

{
  "city": "Los Angeles",
  "city_id": "1"
}


> load csv with headers from "file:///ontime" as ontime create (n:Ontime {city_id:ontime.city_id,	air_line:ontime.air_line,	arrived:ontime.arrived,	n:ontime.n})

> MATCH (n:Ontime) RETURN n LIMIT 1

{
  "arrived": "ontime",
  "n": "497",
  "air_line": "Alaska",
  "city_id": "1"
}
