---
title: "DATA 606 HW3"
author: "Nathan Cooper"
date: "September 26, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DATA606)
```

### 3.2 Area under the curve, Part II.

What percent of a standard normal distributionN($\mu = 0$ , $\sigma = 1$) is found in each region? Be sure to draw a graph.

(a) $Z>-1.13$

```{r a, eval=TRUE}
1- pnorm(-1.13)
normalPlot(bounds = c(-1.13,Inf))
```

(b) $Z<0.18$ 

```{r b, eval=TRUE}
pnorm(0.18)
normalPlot(bounds = c(-Inf,0.18))
```

(c) $Z>8$ 

```{r c, eval=TRUE}
1- pnorm(8)
normalPlot(bounds = c(8,Inf))
```

(d) $|Z| <0.5$

```{r d, eval=TRUE}
pnorm(0.5) -  pnorm(-0.5)
normalPlot(bounds = c(-0.5,0.5))
```

### 3.4 Triathlon times, Part I.
In triathlons, it is common for racers to be placed into age and gender groups. Friends Leo and Mary both completed the Hermosa Beach Triathlon, where Leo competed in the Men, Ages 30 - 34 group while Mary competed in the Women, Ages 25 -29 group. Leo completed the race in 1:22:28 (4948 seconds), while Mary completed the race in 1:31:53 (5513 seconds). Obviously Leo finished faster, but they are curious about how they did within their respective groups. Can you help them? Here is some information on the performance of their groups:

\begin{itemize}
\item The finishing times of the Men, Ages 30 - 34 group has a mean of 4313 seconds with a standard deviation of 583 seconds. 
\item The finishing times of the Women, Ages 25 - 29 group has a mean of 5261 seconds with a standard deviation of 807 seconds. 
\item The distributions of finishing times for both groups are approximately Normal.
\end{itemize}

Remember: a better performance corresponds to a faster finish.

(a) Write down the short-hand for these two normal distributions.

> Men's group, age 30-34 $\mu = 4313 s$ , $\sigma = 583 s$

> Women's group, age 25-29 $\mu = 5216 s$ , $\sigma = 807 s$

(b) What are the Z-scores for Leo's and Mary's finishing times? What do these Z-scores tell you?

> $Z = \frac{x - \mu}{\sigma}$

> Leo's $Z = \frac{4948s-4313s}{583s}$

```{r Z1,eval=TRUE}
(4948-4313)/583
```

> Mary's $Z = \frac{5513s-5261s}{807s}$

```{r Z2,eval=TRUE}
(5513-5261)/807
```

(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.

> The lower the time  the better you did, so you want a small Z-score, perferably negative. In this case Mary did better since her Z is less than Leo's.

(d) What percent of the triathletes did Leo finish faster than in his group?

```{r d2, eval=TRUE}
1- pnorm(1.089194) 
normalPlot(bounds = c(1.089194,Inf))
```

(e) What percent of the triathletes did Mary finish faster than in her group?

```{r e, eval=TRUE}
1- pnorm(0.3122677) 
normalPlot(bounds = c(0.3122677,Inf))
```

(f) If the distributions of finishing times are not nearly normal, would your answers to parts
$(b) - (e)$ change? Explain your reasoning.

> Yes, any outliers or assymetry in the data will change the precentages based on how badly assymetric the data and the number of outliers, high ot low.

### 3.18 Heights of female college students.
Below are heights of 25 female college students.

```{r fhgt, eval=TRUE}
fhgt <- c(54,55,56,56,57,58,58,59,60,60,60,61,61,62,62,63,63,63,64,65,65,67,67,69,73)
fhgt
```

(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information
to determine if the heights approximately follow the 68-95-99.7% Rule.

```{r fhgt_stats, eval=TRUE}
fmn <- mean(fhgt)
fsd <- sd(fhgt)
j = 0
k = 0
l = 0
for(i in fhgt){
  if(i <= (fmn+3*fsd) & i >= (fmn-3*fsd)){
    j = j+1
    if(i <= (fmn+2*fsd) & i >= (fmn-2*fsd)){
      k = k+1
      if(i <= (fmn+fsd) & i >= (fmn-fsd)){
        l = l+1
  }
    }
      }
  }
 print(paste(as.character(j/length(fhgt)),"are within 3 stdev"))
 print(paste(as.character(k/length(fhgt)),"are within 2 stdev"))
 print(paste(as.character(l/length(fhgt)),"are within 1 stdev"))
```

> These data seem to follow the 68-95-99.7% rule well.

(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs
provided below.

```{r qq-fhgt, eval=TRUE}
hist(fhgt, xlab = "Female Height in Inches", main = "Histogram of Female Height")
qqnorm(fhgt)
qqline(fhgt)
qqnormsim(fhgt)
```

> For a small data set it follows a Normal Distribution fairly well. It would be unreasonable to expect sample 25 random people to make a prefect Bell Curve. 

### 3.22 Defective rate.
A machine that produces a special type of transistor (a component of computers) has a 2% defective rate. The production is considered a random process where each transistor is independent of the others.

(a) What is the probability that the 10th transistor produced is the first with a defect?

> We'll treat a defect as a success for mathematical reasons, a defect is the event of interest. 

```{r geo1, eval=TRUE}
n = 10
p= 0.02
q = 0.98
prb = (q^(n-1)*p)
prb
```

> 1.667496%

(b) What is the probability that the machine produces no defective transistors in a batch of 100?

```{r geo2, eval=TRUE}
n = 100
p= 0.02
q = 0.98
prb = (q^(n))
prb
```

> 13.26196% chance of no defects in a batch of 100. 

(c) On average, how many transistors would you expect to be produced before the first with a defect? What is the standard deviation?

> To answer this we need to find the reciprocol of the failure rate of 2%, $\mu = 1/p = 1/0.02 = \textbf{50}$ , $\sigma = \sqrt{q/p^2} = \sqrt{0.98/0.02^2} = \textbf{49.497}$

(d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others. On average how many transistors would you expect to be produced with this machine before the first with a defect? What is the standard deviation?

> To answer this we need to find the reciprocol of the failure rate of 5%, $\mu = 1/p = 1/0.05 = \textbf{20}$ , $\sigma = \sqrt{q/p^2} = \sqrt{0.95/0.05^2} = \textbf{19.494}$

(e) Based on your answers to parts (c) and (d), how does increasing the probability of an event affect the mean and standard deviation of the wait time until success?

> Both the mean, $\mu$, and standard deviation, $\sigma$ decrease as probabilty increases. 


### 3.38 Male children.
While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.

(a) Use the binomial model to calculate the probability that two of them will be boys.

> I use both the r function dbinom and the formula $P  = {{n}\choose{k}} p^k*q^{n-k}$

```{r binom1, eval=TRUE}
dbinom(2,3,0.51)
n = 3
k = 2
p= 0.51
q = 0.49
prb = (factorial(n)/(factorial(k)*factorial(n-k))*p^k*q^(n-k))
prb
```
(b) Write out all possible orderings of 3 children, 2 of whom are boys. Use these scenarios to
calculate the same probability from part (a) but using the addition rule for disjoint outcomes.
Confirm that your answers from parts (a) and (b) match.


$$P(\{f,m,m\} or\{m,f,m\} or \{m,m,f\}) = P(\{f,m,m\}) + P(\{m,f,m\}) + P(\{m,m,f\})$$ 
$$(0.49*0.51*0.51)+(0.51*0.49*0.51)+(0.51*0.51*0.49) = 0.382347$$

(c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have
3 boys, briefly describe why the approach from part (b) would be more tedious than the
approach from part (a).

> With 8 children and 3 boys you would need to figure out how many permutations there are for 3 boys out of 8 children by hand. There are $8!/(8-3)! = 336$ permutations in this case. In using the binomial distribution you just plug into the formula:

```{r binom2, eval=TRUE}
dbinom(3,8,0.51)
n = 8
k = 3
p= 0.51
q = 0.49
prb = (factorial(n)/(factorial(k)*factorial(n-k))*p^k*q^(n-k))
prb
```


### 3.42 Serving in volleyball.
A not-so-skilled volleyball player hasa 15% chance of making the serve, which involves hitting the ball so it passes over the net on a trajectory such that it will land in the opposing team's court. Suppose that her serves are independent of each other.

(a) What is the probability that on the 10th try she will make her 3rd successful serve?

> Use the negative Binomial Distribution. 

```{r neg1, eval=TRUE}
n = 10
k = 3
p= 0.15
q = 0.85
prb = (factorial(n-1)/(factorial(k-1)*factorial(n-k))*p^k*q^(n-k))
prb
```

> 3.895012 %

(b) Suppose she has made two successful serves in nine attempts. What is the probability that
her 10th serve will be successful?

> 15% since the serves are independant.


(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?

> As given the prior history does not matter when making the serves. So the 10th serve had a 15% chance of success regardless of what came before. When you look at all the possible scenerios for 10 shots 15% of them will have the 10th shot be successful. Two previous successes is only one scenerio out of many.