---
title: "DATA 605 HW12"
author: "Nathan Cooper"
date: "November 17, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(suppressWarnings(library('tidyverse')))
```

## 1.  Provide a scatterplot of LifeExp~TotExp, 
and run simple linear regression. Do not transform the 
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss 
whether the assumptions of simple linear regression met.

First we can load it into a data frame:

```{r, who, eval=TRUE}
who <- "who.csv" %>% read.csv(stringsAsFactors = FALSE) %>% data.frame
head(who) # look at the top of the df
str(who) #look at the data stored in the df
pairs(who[,-1], gap = 0.5, col = "orangered") # [,-1] to remove the country name colunm
```

We can now run the linear regression:

```{r lr1, eval=TRUE}
fit1 <- lm(LifeExp ~ TotExp, data = who)
summary(fit1)
plot(who$TotExp, who$LifeExp, xlab = "Total Expenditures ($)" ,ylab = "Life Expectancy (yrs)", col = "steelblue")
abline(fit1, col="yellow3")
hist(resid(fit1), main = "Histogram of Residuals", xlab = "residuals")
plot(fitted(fit1), resid(fit1))
```

The p-value suggests a statistically significant correlation between total expenditures and life expectancy, since $p<<0.05$. The R$^2$ of 0.2577 means that about 25.77% of the variability of life expectancy about the mean is explained by the model. This is a moderately weak correlation. The F-statistic tells us that adding the variable 'total expenditures' to the model improves the model compared to only having an intercept. The residual standard error tells us that, if the residuals are normally distributed, about 64% of the residuals are between $\pm 9.371$ years. These statistics suggest we have a useful model. 

The linear model, when plotted over the data, does not match the data very closely. Furthermore, the residual analysis shows that the residuals have a strong right skew and do not show constant variance. Therefore, the linear model is not valid in this case.

##2.  Raise life expectancy to the 4.6 power 
(i.e., LifeExp^4.6). Raise total expenditures to the 0.06 
power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r 
re-run the simple regression model using the transformed variables. Provide and interpret the F 
statistics, R^2, standard error, and p-values. Which model is "better?"

```{r lr2, eval=TRUE}
le_4.6 <- who$LifeExp^4.6
te_0.06 <- who$TotExp^0.06
fit2 <- lm(le_4.6 ~ te_0.06)
summary(fit2)
plot(who$TotExp^0.06, who$LifeExp^4.6, xlab = "Total Expenditures^0.06 ($^0.06)" ,ylab = "Life Expectancy^4.6 (yrs^0.06)", col = "steelblue")
abline(fit2, col="yellow3")
hist(resid(fit2), main = "Histogram of Residuals", xlab = "residuals")
plot(fitted(fit2), resid(fit2))
```

The p-value suggests a statistically significant correlation between total expenditures^0.06 and life expectancy^4.6, since $p<<0.05$. The R$^2$ of 0.7298 means that about 72.98% of the variability of life expectancy about the mean is explained by the model. This is a moderately strong correlation. The F-statistic tells us that adding the variable 'total expenditures' to the model improves the model compared to only having an intercept 507.7 is much larger than 65.26, so it is a better fit than before. Note that The residual standard error tells us that, if the residuals are normally distributed, about 64% of the residuals are between $\pm 90490000$ years^4.6. These statistics suggest we have a useful model. 

The linear model, when plotted over the data, matches the data more closely. Furthermore, the residual analysis shows that the residuals are normally distributed and show constant variance; there is no noticeable trend. Therefore, the linear model is valid in this case.

This model is better than in part 1.

## 3.  Using the results from 3
forecast life expectancy when TotExp^.06 =1.5. 

Then forecast life 
expectancy when TotExp^.06=2.5. 

$$
y = -736527910 + 620060216x \\
y = -736527910 + 620060216(1.5) \\
y = 193562414 \\
le = y^{1/4.6} \\
le = 193562414^{1/4.6} \\
le = 63.31153 \space years
$$

Life expectancy is about 63.3 years when tot_exp^0.06 = 1.5.

$$
y = -736527910 + 620060216x \\
y = -736527910 + 620060216(2.5) \\
y = 813622630 \\
le = y^{1/4.6} \\
le = 813622630^{1/4.6} \\
le = 86.50645 \space years
$$

Life expectancy is about 86.5 years when tot_exp^0.06 = 2.5.

## 4.  Build the following multiple regression model 
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp
and interpret the F Statistics, R^2, standard error, 
and p-values. How good is the model?

```{r lr3, eval=TRUE}
fit3 <- lm(LifeExp ~ PropMD + TotExp + PropMD*TotExp, data = who)
summary(fit3)
hist(resid(fit3), main = "Histogram of Residuals", xlab = "residuals")
plot(fitted(fit3), resid(fit3))
```

The p-value is < 0.05 so the model is statistically significant. The F-statistic of 34.49 tells us that adding the 3 variables performs better than just the intercept, but barely as the F-statistic penalizes you for adding variables. R$^2$ of 0.3574 means that 35.74% of the variability about the mean of life expectancy is explained by these 3 variables. This is a moderately weak correlation. Residual standard error of 8.765 means that if the residuals are normally distributed, 64% will be $\pm$ 8.765 years.

The residual analysis shows that the residuals have a strong right skew and do not show constant variance. Therefore, the linear model is not valid in this case.

## 5.   Forecast LifeExp 
when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why 
or why not?

$$
LifeExp = 6.277*10^{1} + 1.497*10^{3}*PropMD +7.233*10^{-5}TotExp -6.026*10^{-3}*PropMD*TotExp \\
LifeExp = 6.277*10^{1} + 1.497*10^{3}*0.03 +7.233*10^{-5}*14 -6.026*10^{-3}*0.03*14 \\
LifeExp = 107.6808
$$

107.7 years seems very unrealistic first, that age for a human being is an outlier to the point where someone making it to that age will be featured in National or even International news. Life expectancy is the average time a person can expect to live, I don't see approximately 50% of a population making it to 107 with modern technology. Furthermore, 3% of the population being MDs seems unreasonably high; that would be about 9.5 million doctors in the US. Total Expenditures of $14 seems unreasonably low as that includes both personal and government expenditures. 