---
title: "R_MySQL_MovieDB"
author: "Nathan Cooper"
date: "September 9, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'll use the RMySQL pacakge to connect to MySQL, first we need to install the package and load it.

```{r RMySQL-Load, eval= TRUE}
#install.packages("RMySQL")
library(RMySQL)
```
Now to make the connection. With some tests to make sure the connection works.

```{r db-connect, eval=TRUE}
movie_db = dbConnect(MySQL(), user='*******', password='************', dbname='movies', host='localhost')

dbListTables(movie_db)
dbListFields(movie_db, 'movies')
```

I sourced my data on Facebook using the movies, Star War: The Force Awakens, Star Wars: Rogue One, Deadpool, Trolls, Fantastic Beasts and where to Find Them, and Kubo of the Two Strings. My friends are a mix of geeks and/or parents of young kids so I included movies likely to get a response. I recieved more then required by the assignment within minutes.

now to Read the database into the dataframe:

```{r movie_df,  eval=TRUE}
movie_df <- dbReadTable(movie_db, name = "movies")
movie_df
movie_df$SW_TFA
```

While we are at it let's take a look at the data, keeping in mind that 0 means they did not see the film. Since I did have friends that insisted in half scores, histograms work a little better than bar plots in R even though the data is still discrete.

```{r movie-barplots}
hist(movie_df$SW_TFA)
hist(movie_df$SW_R1)
hist(movie_df$Dpool)
hist(movie_df$Trolls)
hist(movie_df$FanBeast)
hist(movie_df$Kubo)
```

It looks like Deadpool was the most liked movie with Kubo coming in a close second, Trolls was the least liked an Fantastic Beasts had the most spread in opinion.