---
title: "DATA 605 HW7"
author: "Nathan Cooper"
date: "October 12, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Distribution of N Uniform Random Variables
Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi's. Find the distribution of Y.

Drawing from: https://www2.stat.duke.edu/courses/Spring12/sta104.1/Lectures/Lec15.pdf 

Slide 7 gives the main argument to this problem.

and 

https://danieltakeshi.github.io/2016/09/25/the-expectation-of-the-minimum-of-iid-uniform-random-variables/

The beginning part of this page takes the calculation done on slide 7 above and shows how the products  of the CDF come about more clearly.

For an individual $X_i$ the CDF, $F(x)$ where $x \in (1,k)$ is:

$$
F(x) = \int_1^x \frac{dx}{k-1} = \frac{x-1}{k-1}
$$

Note also that Probability, $P(X_i<x) = F(x)$. We will substitute this expression in the final stages of the calculation below. Which $X_i$ is the minimum $minX_i$ is arbitrary, and can be defined by it's compliment:

$$
P(min X_i < x) = 1 - P(X_i \geq x) \\ 
P(min X_i < x) = 1 - \prod_{i=1}^n[1 -P(X_i \leq x)] \\
P(min X_i < x) = 1-[1 - P(X_i \leq x)]^n \\
P(min X_i < x) = 1- [1 - \frac{x-1}{k-1}]^n \\
\bf{F(Y) = 1-[1- \frac{x-1}{k-1}]^n}
$$

Where Y is the distribution of the minimum $X_i$, as defined by the problem text.


##2. Copy Machine Failure Probabilty 
Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer's expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).

If we can expect 1 failure every 10 years this means that the probability (p) for failure in a given year is 0.10 since $E(x) = \sum xm(x) = 10(p) = 1$. We expect 1 failure, there are 10 years and each year has an equal chance of failure. 

### a. Geometric 
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

From above $p = 0.1$, and $q = 1- p =0.9$
According to the hint we assume the copier fails on or after the 9th year so n = 9.

$$
P(X = n) = q^{n-1}*p \\
P(X \leq 9) = \sum_{i=0}^{8}q^i*p = \\
0.1+0.9*0.1+0.9^2*0.1+0.9^3*0.1+...+0.9^80.1=0.6125795
$$
To get the probability that the copier fails after 8 years we take the compliment $P(X \geq 9) = 1- 0.6125795 = 0.3874205$

We can verify using pgeom in r with x = 8 and p = 0.1 and taking the compliment:

```{r geom, eval=TRUE}
1-pgeom(8,0.1)
```

According to the Geometric Distribution there is a 38.74% probability of the copier failing after 8 years.

For the Geometric Distribution the Expected Value is simply $E[X] =\frac{1}{p} = 10\space years$ and the Standard Deviation is $SD[X] = \sqrt{\frac{q}{p^2}} = 9.486833\space years$.

### b. Exponential 
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.

For the exponential density we will let $\lambda = 0.1$ since $\mu = 10$ and $\lambda = \frac{1}{\mu}$.

From the text, the probability that there is a failure within 8 years is:

$$
P(T \leq 8) = \int_0^8 0.1e^{-0.1t}dt = 1-e^{-0.1*8} = 0.550671
$$

The probability of failure after 8 years failure is the compliment $e^{-0.1*8} = 0.449329$ 

```{r exp, eval=TRUE}
1-pexp(8,0.1)
```
According to Exponential Density there is a 44.93% probability of the copier failing after 8 years.

The expected value, taken on the interval $0 to \infty$ is defined as:

$$
E[X] = \int_0^{\infty}xf(x)dx = \int_0^{\infty}\lambda xe^{-\lambda x}dx \\
\int_0^{\infty}\lambda xe^{-\lambda x}dx = \frac{\lambda}{\lambda^2}  = \frac{1}{\lambda}\\
E[X] = \mu  = 10\space years
$$ 

I used wolframalpha.com to evaluate $\int_0^{\infty}\lambda xe^{-\lambda x}dx$

Now for Standard Deviation:
$$
Var[X] = \int_0^{\infty}(x - \mu)^2f(x)dx = \int_0^{\infty}\lambda x^2e^{-\lambda x}dx - \mu^2 \\
\int_0^{\infty}\lambda x^2e^{-\lambda x}dx = \frac{2\lambda}{\lambda^3}  = \frac{2}{\lambda^2}\\
Var[X] = 2\mu^2 - \mu^2  = 100\space years^2 \\
SD[X] = \sqrt{Var[X]} = 10\space years
$$

I used wolframalpha.com to evaluate $\int_0^{\infty}\lambda x^2e^{-\lambda x}dx$

### c. Binomial
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

We can take $p = 0.1$ and $q =0.9$

The probability that there are 0 successes and 8 failures is:

$$
P(X\leq n) = p^0q^n\\
P(X \leq 8) = p^0q^8 = 0.9^8 = 0.4304672
$$

We can verify in r using pbinom with x=0, size = 8, p = 0.1

```{r bimnom, eval=TRUE}
pbinom(0,8,0.1)
```

According to the Binomial Distribution there is a 43.05% chance of failure after 8 years. As there are no failures within 8 years, the copy machine will fail some time after that.

The expected value for a binomial distribution is:

$E[X] = np = 8*0.1 = 0.8\space failures$

And Standard Deviation is:

$SD[X] = \sqrt{Var[X]} = \sqrt{npq} =\sqrt{8*0.1*0.9} = 0.8485281\space failures$

### d. Poisson 
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

$$
P(X=x) = \frac{\lambda^x}{x!}e^{-\lambda}
$$
In this case we let x = 0 (0 copy machine failures) such that:

$$
\lambda^0 = 1 \\
0! =1 \\
P(X=0) = e^{-\lambda}
$$

Where 
$$
\lambda = \frac{np}{t} = 0.8\\
n = 8 \\
p =0.1 \\
t =1 \\
$$ 

t is the time interval of interest, 1 year, n is the total number of time periods of interest, 8 years, and p is the probability of the event happening within a time period, 0.1 or 10%.

Finally,

$P(X=0) = e^{-0.8} = 0.449329$

We can check using ppois with q =0 and lambda =0.8

```{r pois, eval=TRUE}
ppois(0,0.8)
```

According to the Poisson Distribution there is a 44.93% chance of failure after 8 years. As there are no failures within 8 years, the copy machine will fail some time after that.

For the Poisson Distribution the Expected Value is simply $E[X] =\lambda = 0.8\space failures$ and the Standard Deviation is $SD[X] = \sqrt{\lambda} = 0.8944272\space failures$ 
