---
title: "SciMethodHW"
author: "Nathan Cooper"
date: "July 26, 2017"
output:
  html_document: default
  pdf_document: default
---

# Loading the data into R
These data are available at <https://github.com/ncooper76/MSDA/blob/master/GradeComparison2.csv>
Be sure to change the Directory to your local directory with the code below

```{r}
grades <- read.csv("C:/Users/Nate/Documents/DataSet/GradeComparison2.csv")
```
#Attaching data
You can use the attach command so you can use colName instead of data$colName syntax. You need to detach before you
use another data set
```{r}
attach(grades)
names(grades)
```
#Getting rid of NA's
Since one teacher teaches more students than the other on colunm has NA's use this code to git rid of them.
```{r}
A_count =A_count[!is.na(A_count)]
A_count
```
#Analyzing the data
Student's t-test will tell you if two samples are drawn from different populations. It assumes that both samples have the same varience and are normally distributed. Conventionally 95% is set as the cutoff for statistical significance, so p has to be <0.05.
```{r}
var.test(N_count,A_count)
t.test(N_count, A_count, var.equal=TRUE)
```
